Example Manuscript Template for a Data Analysis Project

### 2025-04-24 products/manuscript/Manuscript.qmd
commit to the changes

```{r}
#load packages
library("readxl")
library("readr")
library("dplyr")
library("ggplot2")
library("tidyverse")
library("here")

# follow through with the code folder to see the #variables in the data and how i selected the #variables make this data set. 
#location of the data
data_location <- here::here("data","processed-data","Tb_immune_data.rds")

#loading the data
data<- read_rds(data_location)

#summary of the data
#summary(data)
```

@@

## Description of data and data source
The data we used for this project was got from the TRAC dataverse and is fully accessible to everyone. This is considered a safe place where the the Tuberculosis data from the TRAc project is kept and can be used by anyone who wants contribute to TB research. More details about the data can be found [here]!(https://dataverse.unc.edu/dataset.xhtml?persistentId=doi:10.15139/S3/AYOFEU). The data is publicaly available and can be accessed by anyone.

@@

\textbf{inclusion and exclusion critera}

Our study included patients aged 15 years and older who were experiencing their first episode of tuberculosis. To maintain data integrity, we excluded individuals who were receiving hormonal therapy or using hormonal contraceptives. Additionally, patients with comorbidities such as diabetes mellitus, cancer, or HIV were excluded from the study. Conversely, we applied strict exclusion criteria to ensure a focused research population: all individuals under 15 years of age, those with recurrent TB episodes, patients taking any form of hormonal therapy or contraception, and those diagnosed with diabetes mellitus, cancer, or HIV were not enrolled in the study.

The variables of interest were extracted from the TRAC dataset  and exported to an MS Excel sheet. The outcome variable was CD4+ immune activatin counts.  The primary independent variable was sex (cis-gender) and this was categorized as male and female. The covariates were age at diagnosis,BMI, Fat in Kg, lean body mass in kg(LBM in kg), CD4+ and CD4 immune activation counts. other variables  like fat mass index (FMI) and fat-free mass index (FFMI) were also obtained.

## Schematic of workflow
Chromosomal sex and hormones play a role in the modulation of immune responses to mTB and thereby contribute to the general outcome of disease. The first cells that encounter Mtb are known as Macrophages. These initiate a local inflammatory reaction which may differ between the sexes[@Hertz2018]. This causes the long term containment of TB infection to differ between men and women since their adaptive immune responses are different.
 @tbfig-1

@@

@Table1 shows the summary statistics of our data variables.
```{r}
# Load the table
#table1 <- readRDS(here::here("results", "tables", "TABLE1.rds"))
#| label: table-Table1
#| tbl-cap: "Summary distribution of baseline characteristics among males and females."
#| echo: FALSE

# Display the table (use knitr::kable, gt, or another display function)
knitr::kable(table1)

@@

@@

Given that fat content serves as a contributing factor to immune activation in disease progression, individuals with higher fat percentages typically demonstrate reduced susceptibility to tuberculosis, and vice versa. We analyzed the distribution of fat composition between men and women. Figure @plot2-result presents a boxplot illustrating the distribution of fat in kilograms across male and female participants.
```{r}
#| label: plot2-result
#| fig-cap: "Box plot showing Fat in kg among males and females."
#| echo: FALSE
knitr::include_graphics(here("results","figures","boxplot2.png"))
```

## Full analysis
In order to take care of confounding, we performed a multi-regression analysis and the results of the analysis are summarized in the table below.

>>>>

@@

## Description of data and data source
The data we used for this project was got from the TRAC dataverse and is fully accessible to everyone. This is considered a safe place where the the Tuberculosis data from the TRAc project is kept and can be used by anyone who wants contribute to TB research. More details about the data can be found [here](https://dataverse.unc.edu/dataset.xhtml?persistentId=doi:10.15139/S3/AYOFEU). The data is publicly available and can be accessed by anyone.

@@

###inclusion and exclusion critera

Our study included patients aged 15 years and older experiencing their first episode of tuberculosis. To maintain data integrity, we excluded individuals receiving hormonal therapy or using hormonal contraceptives. Additionally, patients with comorbidities including diabetes mellitus, cancer, or HIV were excluded. Our strict exclusion criteria ensured a focused research population by omitting all individuals under 15 years of age, those with recurrent TB episodes, patients taking any form of hormonal therapy or contraception, and those diagnosed with diabetes mellitus, cancer, or HIV.

The variables of interest were extracted from the TRAC dataset  and exported to an MS Excel sheet. The outcome variable was CD4+ immune activation counts.  The primary independent variable was sex (cis-gender) and this was categorized as male and female. The covariates were age at diagnosis,BMI, Fat in Kg, lean body mass in kg(LBM in kg), CD4+ and CD4 immune activation counts. other variables  like fat mass index (FMI) and fat-free mass index (FFMI) were also obtained.

## Schematic of workflow
Chromosomal sex and hormones play a role in the modulation of immune responses to mTB and thereby contribute to the general outcome of disease. The first cells that encounter Mtb are known as Macrophages. These initiate a local inflammatory reaction which may differ between the sexes[@Hertz2018]. This causes the long term containment of TB infection to differ between men and women since their adaptive immune responses are different. See @tbfig-1

@@

@Table1 shows the summary statistics of our data variables.
```{r}
#| label: tbl-table1
#| tbl-cap: "Summary distribution of baseline characteristics among males and females."
#| echo: FALSE
table1<-readRDS(here::here("results", "tables", "TABLE1.rds"))
table1

@@

Table @tbl-correlation shows the correlation matrix between immune activation markers and body composition variables.

```{r}
#| label: tbl-correlation
#| tbl-cap: "Correlation matrix between immune activation markers and body composition variables."
#| echo: FALSE

# Load the table
cor_table <- readRDS(here::here("results", "tables", "TABLE2_Correlation.rds"))

cor_table

```
Table 2 presents the Pearson correlation coefficients between CD4 immune markers and body composition variables. CD4+ count showed a modest positive correlation with both CD4 immune activation count (r = 0.22) and fat mass (r = 0.17), and a negative correlation with lean body mass (r = –0.18). Notably, fat in kilograms was strongly negatively correlated with lean body mass (r = –0.94), indicating an inverse relationship between fat and lean tissue. Participant age showed very weak correlations with all variables, suggesting age may not substantially confound relationships between immune activation and body composition in this sample.

@@

Given that fat content serves as a contributing factor to immune activation in disease progression, individuals with higher fat percentages typically demonstrate reduced susceptibility to tuberculosis, and vice versa. We analyzed the distribution of fat composition between men and women. This @plot2-result presents a boxplot illustrating the distribution of fat in kilograms across male and female participants.
```{r}
#| label: plot2-result
#| fig-cap: "Box plot showing Fat in kg among males and females."
#| echo: FALSE
knitr::include_graphics(here("results","figures","boxplot2.png"))
```

## Full analysis
In order to take care of confounding, we performed a multi-regression analysis and the results of the analysis were not satifactory.
